Indicators for Ecosystem Condition in Norway

### 2022-10-10 naturtype2.Rmd
nin variables draft

ntypDF2$hovedgruppe[ntypDF2$Nature_type == "Hagemark"                                           ] <- "T32 | special management"

@@

ntypDF2$hovedgruppe[ntypDF2$Nature_type == "Lauveng"                                            ] <- "T32 | special management"

@@

ntypDF2$hovedgruppe[ntypDF2$Nature_type == "Naturbeitemark"                                     ] <- "T32 | special management"

@@

ntypDF2$hovedgruppe[ntypDF2$Nature_type == "Slåttemark"                                         ] <- "T32 | special management"

@@

Cast to get NiN codes as columns
```{r}
ntyp_vars_wide <- data.table::dcast(setDT(ntyp_vars),
                                    Nature_type~NiN_variable_code,
                                    value.var = "percentageUse")
```

This dat set contains the nature types that were only mappen in 2018 or 2019, and which we removed from ntypDF2. 
```{r}
ntyp_vars_wide$Nature_type[ntyp_vars_wide$Nature_type %!in% ntypDF2$Nature_type]
```

@@

@@

>>>>

ntypDF2$hovedgruppe[ntypDF2$Nature_type == "Hagemark"                                           ] <- "T32 | all (if including sub-types)"

@@

ntypDF2$hovedgruppe[ntypDF2$Nature_type == "Lauveng"                                            ] <- "T32 | all (if including sub-types)"

@@

ntypDF2$hovedgruppe[ntypDF2$Nature_type == "Naturbeitemark"                                     ] <- "T32 | all (if including sub-types)"

@@

ntypDF2$hovedgruppe[ntypDF2$Nature_type == "Slåttemark"                                         ] <- "T32 | all (if including sub-types)"

@@

Cast to get NiN codes as columns
```{r}
# should switch to pivot_wider
ntyp_vars_wide <- data.table::dcast(setDT(ntyp_vars),
                                    Nature_type~NiN_variable_code,
                                    value.var = "percentageUse")
```

This data set contains the nature types that were only mapped in 2018 or 2019, and which we removed from ntypDF2. 
```{r}
ntyp_vars_wide$Nature_type[ntyp_vars_wide$Nature_type %!in% ntypDF2$Nature_type]
```

@@

From the figure above I take that *Naturlig åpne områder* is very poorly represented in general, and that there are many NiN main types that are completely missing from the dataset. The three most common types, however, have complete thematic coverage (in excluding the year 2018 and including all sub-types). I will need to investigate if there are some NiN variables for the nature types mapping T18 Åpen flomfastmark, T12 Strandeng and T21 Sanddynemark which we can use.

For *Semi-naturlig mark* we have quite good thematic coverage for the thre main types.The other three types probably makes up a considerably smaller area, but note that T33 has some coverage, and is classes as *all (-2018)*. Of the thre main ecosystems, semi-naturlig mark dominates clearly. I will need to look for NiN variabels associated with the maooing units for these main NiN types. I can also possibly explore the distribution of the mapped polygons in environmental space and compare it agains AR5 as a reference.

For *Våtmark* we have good thematic coverage for V3 nedbørsmyr and V9 semi-naturlig myr, but for V1 åpen jordvassmyr only calcareous localities are mapped and assessed. A question is then whether, for a given NiN variable, that calcareous localities can be representative for all mires (ppor and rich) or if this will intrucude too much bias in one way or another.

I will give priority to *Naturlig åpne områder* because of the poor data availabiliy for this ecosystem in general.

@@

```{r}
areaSum <- ntyp_fill2 %>%
  group_by(Ecosystem) %>%
  summarise(km2 = sum(km2)) %>%
  tibble::add_column(.before=1,
                     NiN_code = rep("Combined", 3))

```{r, fig.cap="Barplot showing the proportion area for which each NiN variable is recorded. The last letter of the NiN variable code refers to the ecosystem tye, and is not part of the variable name. The total mapped area for each main ecosystem (the facets) is shown as a dashed red line. "}
ntyp_fill2 %>%
  pivot_longer(cols = unique(ntyp_vars$NiN_variable_code)) %>%
  drop_na(value) %>%
  rename(NiN_code = name,
         useFrequency = value) %>%
  filter(useFrequency > 0) %>%
  group_by(NiN_code, Ecosystem) %>%
  summarise(km2 = sum(km2)) %>%
  arrange(km2) %>%
  mutate(lab = paste(NiN_code, substr(Ecosystem, start = 1, stop = 1)), sep = " ") %>%
  group_by(Ecosystem) %>%
  mutate(lab = forcats::fct_reorder(lab, km2)) %>%
  ggplot(aes(x = lab,
              y = km2))+
  geom_bar(stat = "identity",
           colour = "grey40",
           fill = "grey80")+
  geom_hline(data = areaSum,
             aes(yintercept = km2),
             colour = "red",
             linetype = "dashed")+
  facet_wrap(vars(Ecosystem),
             scales = "free",
             labeller = label_wrap_gen(width=25))+
  labs(x = "",
       y = expression(km^2))+
  coord_flip()+
  theme_bw(base_size = mySize)
```
